---
title: "sewanee_climate"
author: "Jawaria Jaleel"
output: "html_document"
date: "2024-06-12"
---
#Introduction

### Questions

#1. Average precipitation for each month in five year periods.
```{r}

```


#2. Effect of temperature on precipitation from 2010 to 2020


#3. Minimum and maximum temperature over time for the entire period


#4. How has snowfall changed over time for the entire period?


#5. Average temperature over time.


#6. What month has the most snowfall on average
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

#load library
library(tidyverse)
setwd("/Users/jawariajaleel/Desktop/dl_2024/climate_sewanee/")
weather <- read_csv("/Users/jawariajaleel/Desktop/dl_2024/climate_sewanee/sewanee_weather.csv")
# weather <- read_csv("sewanee_weather.csv")
#make a col for month and year
weather <- weather %>%
  mutate(month=month(ymd(DATE))) %>%
  mutate(year=year(ymd(DATE))) %>%
  mutate(month_name = month.abb[month]) %>%
  mutate(month_name=fct_reorder(month_name, month))

#finding monthly avg for five year periods
weather <- weather %>%
  mutate(yrblock = case_when(
      year >= 2000 & year <= 2004  ~ "2000-2004",
      year >= 2005 & year <= 2009  ~ "2005-2009",
      year >= 2010 & year <= 2014  ~ "2010-2014",
      year >= 2015 & year <= 2019  ~ "2015-2019",
      year >= 2020 & year <= 2024  ~ "2020-2024",
      .default = "other"
    ))

```

```{r}
#c#create a new datasets with months, year blocks, and average precipitation
precip <- weather %>%
  group_by(yrblock, month_name) %>%
  summarize(meanprecip = mean(PRCP, na.rm = TRUE))
```
```{r}
#Graphs
ggplot(precip, aes(x = month_name, y=meanprecip))+
  geom_col(fill="lightblue")+
  facet_wrap(~yrblock, nrow = 5)
  labs(x = "Month", y = "Average Precipitation")
  
```


```{r}
# Average precipication and average temperature
```{r}
temp <- weather %>%
  group_by(yrblock, month_name) %>%
  summarize(meantemp = mean(TMAX, na.rm = TRUE))
```
```{r}
ggplot(temp, aes(x = month_name, y=meantemp))+
  geom_col(fill="#FF7074")+
  facet_wrap(~yrblock, nrow = 5)
  labs(x = "Month", y = "Average Temperature")
```
```{r}
tempmin <- weather %>%
  group_by(yrblock, month_name) %>%
  summarize(meantemp = mean(TMIN, na.rm = TRUE))
```
```{r}
ggplot(tempmin, aes(x = month_name, y=meantemp))+
  geom_col(fill="pink")+
  facet_wrap(~yrblock, nrow = 5)
  labs(x = "Month", y = "Average Temperature")
```












```

```